\chapter[SOLVING WORD PROBLEMS]{SOLVING WORD PROBLEMS\\ IN GEOMETRY}
\section*{INTRODUCTION}
This module is intended to give an overview on solving word problems in geometry. It presents
the different steps in problem solving as designed by Polya. It also contains sample word problems for
the participants to solve. Assessment questions are also given.
\section*{OBJECTIVES}
After completion of the module, the participants are expected to:
\begin{enumerate}
\item Apply the different steps in problems solving
\item Determine the appropriate strategy to use in solving a word problem.
\item Use variety of problem solving strategies in solving word problems.
\item Solve the word problems accurately.
\item Appreciate the importance of problem solving in a mathematics curriculum.
\item Gain confidence in doing problem solving
\item Create more interesting word problems.
\item Integrate or connect geometry to origami, the art of paper folding.
\end{enumerate}
\section*{DISCUSSION}
The National Council of Teachers of Mathematics (NCTM) gives the following definitions on the
nature of mathematics: mathematics is a study of patterns and relationships, it is a way of thinking, it is
an art, it is a language and it is a tool. NCTM identified five broad goals required to meet the students’
mathematical needs for the 21st century. They are as follows: students should (1) value mathematics,
(2) reason mathematically, (3) communicate mathematics, (4) solve problems and (5) develop
confidence.

Problem solving is the heart of mathematics. It is a process. It is the means by which an
individual uses previously acquired knowledge, skills and understanding to satisfy the demands of an
unfamiliar situation. Students must be exposed to a variety of problems – problems that vary in context,
in level of difficulty and in methods in solving the problem. But what is really a problem or a word
problem specifically? A word problem is a mathematical proposition that a student has never
experienced solving. What are the purposes of word problems? The following are the different
purposes of word problems: (1) to prepare the students for real life, (2) to develop children’s logical and
abstract thinking and mental discipline, and (3) to motivate students. Hence a word problem should
simulate real life situations, events, and places that students can experience firsthand. In this way the
students will be able to see the importance of mathematics in their life.

George Polya, a mathematician, devised the following \Bold{four-step method} in solving word problems:

\begin{enumerate}
\item \textbf{Read and understand the problem}\index{four-step method!Read and understand the problem}

Read and reread the problem. Ask the following questions: What is the situation all
about? What information is given? What are the different assumptions? What information is
missing? What are you being asked to find or to do? You can make a list in this format

Information given (GIVEN): \makebox[2in]{\hrulefill}

Goals of the problem (REQUIRED): \makebox[2in]{\hrulefill}

\item \textbf{Plan how to solve the problem. Try a strategy}\index{four-step method!Try a strategy}\index{four-step method!Plan how to solve the problem}

In developing a plan to solve the problem ask the following questions: Have you ever
worked a similar problem before? Will you estimate or calculate? What strategy (ies) can you
use? Consider the different strategies in solving problem that you know. The following could be
your shopping list on the different problem solving strategies: draw a diagram, make a table,
write an equation, guess and check (trial and error), look for a pattern, solve a simpler problem,
work backward. Word problems can be solved using different strategies.

\item \textbf{Solve the problem/ Carry out the plan.}\index{four-step method!solve the problem}\index{four-step method!carry out the plan}

As you implement the strategy that you choose, ask the following questions: Did you
calculate correctly? What is the solution? Did you interpret correctly? Did you answer the
question?

\item \textbf{Look back}\index{four-step method!look back}

Most students forget this last part especially if they are solving a problem set or many
problems. The most important question that should be asked here is: “Is the calculated answer
reasonable?”
\end{enumerate}
However there is also a danger in over-using Polya's four steps in problem solving. Students
may think that problem solving is one directional that follows the given steps above as if it's the
prescribed recipe. Emphasize to students that with challenging problems, the actual problem solving
becomes a process whereby the solver keeps a mental "check" of the progress, and corrects himself if
progress is not made. He may go one route, notice it won't work, go backwards a bit, and then, take
another route. In other words, devising plans and carrying them out can occur somewhat
simultaneously, and we can go back and forth. Lastly, not all problems can be solved using the steps
above.

\subsection*{Problems Involving Perimeters}
\begin{example}
\Item The perimeter of a square is 28 cm. Find the length of each side.

\Solution

\textit{Geometry Concept:} A square is a quadrilateral (4-sided polygon) with four congruent sides. Its four
interior angles are all right angles. The formula for the perimeter ($P$) of a square is $P = 4s$ and its area ($A$) is $A = s^2$ where $s$ is the side of the square.

\begin{tikzpicture}
\draw (0,0) -- (1,0) -- (1,1) -- (0,1) -- cycle;
\node [below] at (0.5,0) {$s$};
\node [right] at (1,0.5) {$s$};
\node [above] at (0.5,1) {$s$};
\node [left] at (0,0.5) {$s$};
\end{tikzpicture}
\begin{align*}
P&=4s\\
28&=4s\\
\therefore s&=7
\end{align*}
Therefore, the square has side of length 7 cm.
\end{example}

\begin{example}
\item The length of a rectangular garden is 3 meters more than its width. If the perimeter of the garden is
50 meters, what is the area of the rectangle?

\Solution

\textit{Geometry Concept}

A rectangle is a quadrilateral (4-sided polygon) with two pairs of congruent
and parallel sides. Its four interior angles are all right angles.
The formula for the perimeter ($P$) of a rectangle is $P = 2L + 2W$ where $L$ is
the length and $W$ is the width of the rectangle.

\begin{tikzpicture}[xscale=2]
\draw (0,0) -- (1,0) -- (1,1) -- (0,1) -- cycle;
\node [below] at (0.5,0) {$L$};
\node [right] at (1,0.5) {$W$};
\node [above] at (0.5,1) {$L$};
\node [left] at (0,0.5) {$W$};
\end{tikzpicture}

\textit{Representation}
\begin{align*}
W&=\text{width of the rectangle}\\
L&=W+3
\end{align*}

\textit{Equation:}
\begin{align*}
50 &= 2(W + 3) + 2W && \text{Using the formula for perimeter.}\\
50 &= 2W + 6 + 2W \\
44 &= 4W \\
W &= 11 \\
L &= 11 + 3 = 14 \\
A &= (14)(11) = 154 && \text{Using the formula for area.}
\end{align*}

Therefore, the rectangle has an area of 154 $\cm^2$.
\end{example}

\begin{example}
\Item The longest side of a triangle is twice as long as the shortest side and is also 2 cm longer than the third side. If the perimeter of the triangle is 33 cm, what is the length of each side?

\Solution

\textit{Geometry Concept:}

A triangle is a 3-sided polygon. The formula for the perimeter ($P$) of a triangle is
$P = a + b + c$ where $a$, $b$ and $c$ are the sides of the triangle.

\begin{tikzpicture}
\draw (0,0) -- (2,0) -- (0.5,1.5) -- cycle;
\node [below] at (1,0) {$c$};
\node [right] at (1.25,0.75) {$b$};
\node [left] at (0.25,0.75) {$a$};
\end{tikzpicture}

\textit{Representation}

\begin{align*}
a&=x && \text{shortest side of the triangle}\\
b&=2x && \text{longest side}\\
c&=2x-2 && \text{third side}
\end{align*}

\textit{Equation}

\begin{align*}
33 &= x + 2x + 2x - 2 && \text{(Using the formula for the perimeter.)}\\
35 &= 5x\\
x &=7\\
a&=x=7\\
b&= 2x = 14\\
c &= 2x - 2 = 12
\end{align*}
Therefore, the sides of the triangle are 7 cm, 12 cm and 14 cm.
\end{example}

\begin{example}
\Item What is the measure of the base of an isosceles triangle if its perimeter is 50 cm and the length of one of its congruent sides exceeds the length of the base by 10 cm?

\begin{tikzpicture}[yscale=2]
\draw (0,0) -- (1,0) -- (0.5,1) -- cycle;
\node [below] at (0.5,0) {base};
\node [right] at (0.75,0.5) {leg};
\node [left] at (0.25,0.5) {leg};
\end{tikzpicture}

\Solution

\textit{Geometry Concept}

An isosceles triangle is a triangle with two congruent sides. The congruent sides are
called legs while the non-congruent side is called the base.

\textit{Representation}

\begin{align*}
x&=\text{length of the base of the triangle}\\
  x+10&=\text{length of the leg of the triangle}
\end{align*}

\textit{Equation}

\begin{align*}
50 &= x + 2(x + 10) && \text{(Using the formula for perimeter of triangles.)}\\
50 &= x + 2x + 20\\
30 &= 3x\\
x &= 10
\end{align*}
Therefore, the base of the isosceles triangle is 10 cm.
\end{example}

\subsection*{PROBLEMS INVOLVING ANGLES OF POLYGONS}
\begin{example}
\Item The first angle of a triangle is triple that of the smallest angle. The second angle is 5 degrees more than the first angle. What is the measure of the smallest angle?

\Solution

\textit{Geometry Concept}

A triangle has three interior angles and their measures sum up to 180 degrees.

\textit{Representation}
\begin{align*}
x &= \text{smallest angle of the triangle}\\
3x &= \text{first angle of the triangle}\\
3x + 5 &= \text{second angle of the triangle}
\end{align*}
\textit{Equation}
\begin{align*}
x + 3x + 3x + 5 &= 180 && \text{(Sum of the angles of a triangle is $180^{\circ}$.)}\\
7x &= 175\\
x &= 25
\end{align*}
Therefore, the measure of the smallest angle of the triangle is 25 degrees.
\end{example}

\begin{example}
\Item One of the base angles of an isosceles triangle measures 15 degrees less than its vertex angle. Find the
measure of its vertex angle.

\textit{Geometry Concept}

In an isosceles triangle, it has two congruent angles opposite the congruent sides. The angle opposite
the base is called the vertex angle.

\Solution

\textit{Representation}

\begin{align*}
x&=\text{vertex angle}\\
x-15 &= \text{base angle}
\end{align*}

\textit{Equation}

\begin{align*}
2(x - 15) + x &= 180 && \text{(Sum of the angles of a triangle is $180^{\circ}$.)}\\
2x - 30 + x &= 180\\
3x &= 210\\
x &= 70
\end{align*}
Therefore, the measure of the vertex angle of the triangle is 70 degrees.
\end{example}

\begin{example}
\Item Given the parallelogram on the right, find the value of $x$.

\begin{tikzpicture}
\node (trap1) [trapezium,trapezium left angle=81,trapezium right angle=99,minimum width=3cm, minimum height=1.5cm,draw] {};
\node [below right] at (trap1.north west) {$(6x-15)^{\circ}$};
\node [above right] at ([xshift=-10pt]trap1.south west) {$(4x+5)^{\circ}$};
\end{tikzpicture}

\textit{Geometry Concept}

A parallelogram is a quadrilateral with two pairs of congruent and parallel sides. Its two pairs of opposite
angles are congruent and the sum of the measures of any two consecutive angles is 180 degrees.

\Solution

\begin{align*}
(4x + 5) + (6x - 15) &= 180 && \text{(The given angles are consecutive.)}\\
10x - 10 &= 180\\
10x &= 190\\
x &= 19
\end{align*}
\end{example}

\begin{example}
\Item Given the rhombus, find the measure of the angles of the rhombus.

\begin{tikzpicture}
\node (trap1) [trapezium,trapezium left angle=45,trapezium right angle=135,minimum width=3cm, minimum height=1.5cm,draw] {};
\node [below right] at ([xshift=-15pt]trap1.north east) {$(6x-15)^{\circ}$};
\node [above left] at ([xshift=15pt]trap1.south west) {$(4x+5)^{\circ}$};
\end{tikzpicture}

\Solution

\textit{Geometry Concept}

A rhombus is a parallelogram with four congruent sides. Just like a parallelogram, its two pairs of
opposite angles are congruent and the sum of the measures of any two consecutive angles is 180
degrees.
%
\begin{align*}
6x - 15 &= 4x + 5 && \text{(The given angles are opposite.)}\\
2x &= 20\\
x &= 10\\
4x + 5 &= 45
\end{align*}
The given opposite angles have a measure of 45 degrees each while its other two angles
measure $180 - 45 = 135$ degrees each.

Therefore, the measures of the angles of the rhombus are $45\degree$, $45\degree$, $135\degree$, and $135\degree$.
\end{example}
\subsection*{PROBLEMS INVOLVING COMPLEMENTARY, SUPPLEMENTARY AND CONGRUENT ANGLES}
\begin{example}
\item Given two intersecting lines, find the value of $x$.

\begin{tikzpicture}[scale=2,thick,>=stealth']
\draw [->] (0,0) -- (22.5:1cm);
\draw [->] (0,0) -- (22.5:-1cm);
\draw [->] (0,0) -- (-22.5:1cm);
\draw [->] (0,0) -- (-22.5:-1cm);
\node [right] at ([xshift=5pt]0,0) {$(6x-15)\degree$};
\node [left] at ([xshift=-5pt]0,0) {$(4x+5)\degree$};
\end{tikzpicture}

\Solution

\textit{Geometry Concept}

The two intersecting lines formed angles and the two given angles are “opposite” each other. They are
called \Bold{vertical angles}. Vertical angles are congruent. 

Since the given angles are vertical angle, they are congruent. Refer to Example 8.

Therefore, the value of $x$ is 10.
\end{example}

\begin{example}
\Item Given two intersecting lines, find the value of $x$.

\begin{tikzpicture}[scale=2,thick,>=stealth']
\draw [->] (0,0) -- (22.5:1cm);
\draw [->] (0,0) -- (22.5:-1cm);
\draw [->] (0,0) -- (-22.5:1cm);
\draw [->] (0,0) -- (-22.5:-1cm);
\node [right] at ([xshift=5pt]0,0) {$(6x-15)\degree$};
\node [below] at ([yshift=-3pt]0,0) {$(4x+5)\degree$};
\end{tikzpicture}

\Solution

\textit{Geometry Concept}

The two intersecting lines formed angles and note that when the measures of the two given angles are
added, we get the measure of a straight angle. These two angles are called linear pairs. The degree
measures of a \Bold{linear pair} sum up to 180 degrees.

Since the given angles are linear pairs, they sum up to $180\degree$. Refer to Example 7. %Fix reference here.

Therefore, the value of $x$ is 19.
\end{example}

\begin{example}
\Item The supplement of an angle is 20 degrees more than thrice the measure of the angle. What is the
measure of the angle?

\Solution

\textit{Geometry Concept:} When the measures of two angles sum up to 180 degrees, they are called \Bold{supplementary angles}.
\begin{align*}
x &= \text{the given angle}\\
180 - x &= \text{the measure of the supplement of the angle}\\
180 - x &= 3x + 20\\
160 &= 4x\\
x &= 40
\end{align*}
Therefore, the measure of the angle is 40 degrees.
\end{example}

\begin{example}
\Item The supplement of a certain angle is four times its complement. What is the measure of the angle?

\Solution

\textit{Geometry Concept:} When the measures of two angles sum up to 90 degrees, they are called \Bold{complementary angles}.
\begin{align*}
x &= \text{the given angle}\\
180 - x &= \text{the measure of the supplement of the angle}\\
90 - x &= \text{the measure of the complement of the angle}\\
180 - x &= 4(90 - x)\\
180 - x &= 360 - 4x\\
3x &= 180\\
x &= 60
\end{align*}
Therefore, the measure of the angle is 60 degrees.
\end{example}
